---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


# forestinventory

The R-package `forestinventory` addresses the current interest of combining existing 
forest inventory data, which are derived by field surveys, with additional information
sources such as remote sensing data. The major benefit of these so-called *multisource
inventory methods* is the potential increase of estimation precision without an increase 
in the number of expensive field surveys. Additionally, it also allows for deriving 
estimates of sufficient accuracy for spatial units where terrestrial information 
is scarcely available if not absent.

The aim of `forestinventory` is to facilitate the application of 
multiphase forest inventories by providing an extensive set of functions 
for global and small-area estimation procedures.
The implementation includes all estimators for simple and cluster sampling published 
by Daniel Mandallaz between 2007 and 2014, providing point estimates,
their external- and design-based variances as well as confidence
intervals. The procedures have also been optimized for the use of remote sensing
data as auxiliary information.


# Quick demo

We look at the example dataset `grisons` which comes with our package:
```{r,eval=T, echo = F}
library(forestinventory)
```
```{r, eval=F}
library(forestinventory)
?grisons
```

As the help tells us, `grisons` contains the data of a twophase inventory: We are
provided with LiDAR canopy height metrics at 306 inventory locations, and at 67 subsamples
we have the terrestrially measured timber volume values. We now want to estimate the timber volume
in m^3^/ha within four subdomains A, B, C and D (*small areas*).

If we only use the terrestrial information within the small areas, we call the `onephase`-function:
```{r}
op <- onephase(formula = tvol~1,
               data = grisons,
               phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                 area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D")))

summary(op)
```


We now try to increase the precision of our estimates by applying a *twophase 
estimation method*, where we use the large sample of LiDAR-metrics and 
a linear regression model to specify the relationship
between the remote sensing derived predictor variables and the terrestrial timber volume:
```{r}
sae.2p.uv<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                     small_area = list(sa.col = "smallarea", areas = c("A", "B","C", "D"),
                                       unbiased = TRUE))

summary(sae.2p.uv)
```

We now want to compare the results and performances of the onephase and twophase method. For such issues, the package provides
the `estTable()` function that concatenates the results from the different methods in one `list`:

```{r}
sae.table<- estTable(est.list = list(op, sae.2p.uv), sae = TRUE)

data.frame(sae.table[c(1:6,9)])
```

We can already see that the estimation errors of the twophase estimation are up to 5% smaller than the onephase errors.

The function `mphase.gain()` can now be used to further compare the performance of the methods:

```{r}
mphase.gain(sae.table)
```

The column `gain` tells us that the twophase estimation procedure here leads to a 67.9 % reduction in variance compared to the one-
phase procedure". The column `rel.eff` speciﬁes the relative efﬁciency that can be interpreted as the relative sample size of the one-phase estimator needed
to achieve the variance of the multi-phase (here twophase) estimator. For small area "B" we can thus see that we would
have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same
estimation precision as the twophase extended psynth estimator. 


So in our short example, we were able to considerably improve the estimation precision
when combining the terrestrial data with the remote sensing data. The package `forestinventory`
offers further estimators that can be applied to a wide range of multiphase inventory scenarios.


# Installation

The package can be installed from CRAN:
```{r, eval=F}
install.packages("forestinventory")
```

