---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


# forestinventory

The R-package `forestinventory` adresses the current interest of combining existing 
forest inventory data, which are derived by field surveys, with additional information
sources such as remote sensing data. The major benefit of these so-called *multisource
inventory methods* is the potential increase of estimation precision without an increase 
in the number of expensive field surveys. Additionally, it also allows for deriving 
estimates of sufficient accuracy for spatial units where terrestrial information 
is scarcely available if not absent.

The aim of `forestinventory` is to facilitate the application of 
multiphase forest inventories by providing an extensive set of functions 
for global and small-area estimation procedures.
The implementation includes all estimators for simple and cluster sampling published 
by Daniel Mandallaz between 2007 and 2014, providing point estimates,
their external- and design-based variances as well as confidence
intervals. The procedures have also been optimized for the use of remote sensing
data as auxiliary information.


# Quick demo

We look at the example dataset `grisons` which comes with our package:
```{r,eval=T, echo = F}
library(forestinventory)
```
```{r, eval=F}
library(forestinventory)
?grisons
```

As the help tells us, `grisons` contains the data of a twophase inventory: We are
provided with LiDAR canopy height metrics at 306 inventory locations, and at 67 subsamples
we have the terrestrially measured timber volume values. We now want to estimate the timber volume
in m^3^/ha within four subdomains A, B, C and D (*small areas*).

If we only use the terrestrial information within the small areas, we call the `onephase`-function:
```{r}
op <- onephase(formula = tvol~1,
               data = grisons,
               phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                 area = list(sa.col = "smallarea", areas = c("A", "B", "C", "D")))

summary(op)
```

We compute the estimation error in % as the *Standard Deviation / Point Estimate*:
```{r}
ee.o<- 100*(sqrt(op$estimation$variance) / op$estimation$estimate)
names(ee.o)<- op$estimation$area
ee.o
```

We now try to increase the precision of our estimates by applying a *twophase 
estimation method*, where we use the large sample of LiDAR-metrics and 
a linear regression model to specify the relationship
between the remote sensing derived predictor variables and the terrestrial timber volume:
```{r}
sae.2p.uv<- twophase(formula = tvol ~ mean + stddev + max + q75, data = grisons,
                     phase_id = list(phase.col = "phase_id_2p", terrgrid.id = 2),
                     small_area = list(sa.col = "smallarea", areas = c("A", "B","C", "D"),
                                       unbiased = TRUE))

summary(sae.2p.uv)
```

We again compute the estimation error in % and see how the efficiency has been increased from
the onephase to the twophase approach:
```{r}
ee.uv<- 100*(sqrt(sae.2p.uv$estimation$g_variance) / sae.2p.uv$estimation$estimate)
names(ee.uv)<- op$estimation$area
ee.uv

efficiency<- 100* ((ee.o - ee.uv) / ee.o)
names(efficiency)<- op$estimation$area
efficiency
```

In our short example, we were able to reduce the estimation error by 9 % up to 37 %
when combining the terrestrial data with the remote sensing data. The package `forestinventory`
offers further estimators that can be applied to a wide range of multiphase inventory scenarios.


# Installation

The package will soon be available from CRAN:
```{r, eval=F}
install.packages("forestinventory")
```


# Coming soon ...
A vignette will be released as soon as possible, giving detailed information about the implemented
estimation methods and estimators and how to apply them to inventory scenarios in practice.


